Using numerical modelling to investigate temperature and wave exposure as drivers

### 2019-11-29 morphology_paper.Rmd
checked consistancy of in text figure references. Significance levels on graphs adjusted.

---
title: Using numerical modelling to investigate temperature and wave exposure as drivers
  of macroalgal morphological characteristics
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
bibliography: references.bib
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
  word_document:
    reference_docx: frontiers_template.docx
geometry: margin=1in
fontsize: 10pt
papersize: A4
tables: yes
abstract: 'Macroalgal morphological variation is determined to a large extent by
  a combination of environmental factors, with wave exposure and temperature
  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.                Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
  in overall size and strength increasing traits. In terms of temperature, warmer
  environments have been shown to reduce the overall size of resident and transplanted
  species. However, none of the past studies have identified specific wave and temperature
  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
  has often used simple or two-dimensional models of wave exposure, which do not
  take into account important aspects of the nearshore environment such as wave breaking,
  refraction and diffraction. Furthermore, past studies have often used satellite-derived
  datasets as sources for temperature data; however, such data have been shown to have large bias when 
  applied to the nearshore environment. This study used *in situ* temperature data and wave power
  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*. The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
  Furthermore, the results show differences in wave metrics between species; and between
  deep and shallow populations.'
---

@@

The mean annual coastal water temperature for the study sites located on the western side of the peninsula ranged from 14.5 ± 0.9 ℃ (mean ± SD) at St. Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula (see figure \ref{fig:Temperature variables plot}A). For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study (see figure \ref{fig:Temperature variables plot}A). Oudekraal had the lowest annual mean coastal water temperature of 12.3 ± 0.9. Annual maximum coastal temperatures for study sites located on the western side of the peninsula ranged from 16.3 ℃ at St. Helena Bay to 15.1 ℃ at Olifantsbos. For sites located east of the peninsula, annual maximum coastal water temperatures ranged from 16 ℃ at Buffelsbaai to 16.5 ℃ at Betty’s Bay. Amongst all sites, Kalk Bay had the highest annual maximum temperature of 19 ℃. Annual minimum coastal water temperatures for the study sites on the western side of the peninsula ranged from 12.9 ℃ at St. Helena Bay to 14.1 ℃ at Olifantsbos, and on the western side of the peninsula ranged from 15.0 ℃ at Buffelsbaai to 13.6 ℃ at Betty’s Bay.  The annual range in temperature on the western side on the peninsula was 3.4 at St. Helena Bay to 0.9 at Olifantsbos, and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to 2.9 at Betty’s Bay.

@@

temp_boxplot <- ggplot(data = temp_ann_meta, aes(x = region, y = mean_temp, fill  = region)) + 
  geom_boxplot(show.legend = FALSE, width = 0.5) + 
  stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.format") +
  xlab("") + 
  ylab("") +
  coord_flip() + 
  theme_bw() +
  scale_fill_grey(start = 0.4, end = 0.8) + 
  theme(axis.text.y=element_text(vjust=0.3)) 
```

@@

In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A). This is verified by \ref{fig:Temperature variables plot}B, which shows a significantly lower (p < 0.05, figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.

@@

Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual scale (see figure \ref{fig:wave variable plot}). Mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St. Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (see figure \ref{fig:wave variable plot}A). The maximum significant wave height ranged from 1.4 m at St. Helena Bay to 4.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side ranged from 2.6 m at Buffelsbaai to 4.2 m at Betty’s Bay. Mean peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St. Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay (see figure \ref{fig:wave variable plot}B). The data show no trend in mean peak period for the coastline. These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side, with Miller’s Point showing the highest variation across timescales. Maximum peak period ranged from 18.9 s at St. Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula. When wave variables were grouped by region significant differences were found for both mean significant swell height and mean peak period. The western region experiences a significantly higher mean significant swell height (p < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (*p* < 0.05; see figure \ref{fig:wave variable plot}D).

@@

wave_boxplot_1 <- ggplot(data = wave_ann_meta, aes(x = region, y = hs_mean, fill  = region)) + 
  geom_boxplot(show.legend = FALSE) + 
  stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.format") + 
  xlab("") + 
  ylab("") + 
  coord_flip() + 
  theme_bw() +
  scale_fill_grey(start = 0.4, end = 0.8) + 
  theme(axis.text.y=element_text(vjust=0.3))

wave_boxplot_2 <- ggplot(data = wave_ann_meta, aes(x = region, y = tp_mean, fill  = region)) + 
  geom_boxplot(show.legend = FALSE) + 
  stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.format") + 
  xlab("") + 
  ylab("") +
  ylim(8,12) + 
  coord_flip() + 
  theme_bw() +
  scale_fill_grey(start = 0.4, end = 0.8) + 
  theme(axis.text.y=element_text(vjust=0.3))

@@

In Figure \ref{fig:wave power map} the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length). The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits. The western periphery of the Cape Peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see figure \ref{fig:wave power map}). There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.

@@

#temperature and lamanaria  
LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.62, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina weight") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina length") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina thickness") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe length (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe diameter (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("No. of digits") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Thallus mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.format", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Total length (cm)") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

@@

LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.85, label.y.npc = c('top')) + 
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("Lamina length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("Lamina thickness") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.format", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe diameter (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

>>>>

---
title: Using numerical modelling to investigate temperature and wave exposure as drivers
  of macroalgal morphological characteristics
author: "R. Coppin, C. Rautenbach, T. Ponton and A.J Smit"
output:
  word_document:
    reference_docx: frontiers_template.docx
  pdf_document:
    fig_caption: yes
  html_document:
    df_print: paged
bibliography: references.bib
papersize: A4
tables: yes
abstract: 'Macroalgal morphological variation is determined to a large extent by
  a combination of environmental factors, with wave exposure and temperature
  perhaps the main influences as they are key environmental properties to which a species becomes locally adapted.                Macroalgae have shown to exhibit different responses to different magnitudes of exposure to waves, such as reduction
  in overall size and strength increasing traits. In terms of temperature, warmer
  environments have been shown to reduce the overall size of resident and transplanted
  species. However, none of the past studies have identified specific wave and temperature
  metrics responsible for the morphological adaptation macroalgae exhibit. Past research
  has often used simple or two-dimensional models of wave exposure, which do not
  take into account important aspects of the nearshore environment such as wave breaking,
  refraction and diffraction. Furthermore, past studies have often used satellite-derived
  datasets as sources for temperature data; however, such data have been shown to have large bias when 
  applied to the nearshore environment. This study used *in situ* temperature data and wave power
  metrics calculated from a 3D-numerical model to identify specific temperature and wave metrics
  responsible for morphological adaptation of the kelp, *Ecklonia maxima* and *Laminaria pallida*. The results identify wave      exposure as the main driver of morphological adaptation while identifying specific wave metrics.
  Furthermore, the results show differences in wave metrics between species; and between
  deep and shallow populations.'
---

@@

The mean annual coastal water temperature for the study sites located on the western side of the peninsula ranged from 14.5 ± 0.9 ℃ (mean ± SD) at St. Helena Bay, the most northern site on the western side of the peninsula, to 14.6 ± 0.3 ℃ at Olifantsbos, the most southern site on the western side of the peninsula (see figure \ref{fig:Temperature variables plot}A). For sites located on the eastern side of the peninsula (within False Bay), the annual mean coastal water temperatures ranged from 15.5 ± 0.9 ℃ at Buffelsbaai to 15.0 ± 0.9 ℃ at Betty’s Bay, the most eastern site on the coastline in this study (see figure\ref{fig:Temperature variables plot}A). Oudekraal had the lowest annual mean coastal water temperature of 12.3 ± 0.9. Annual maximum coastal temperatures for study sites located on the western side of the peninsula ranged from 16.3 ℃ at St. Helena Bay to 15.1 ℃ at Olifantsbos. For sites located east of the peninsula, annual maximum coastal water temperatures ranged from 16 ℃ at Buffelsbaai to 16.5 ℃ at Betty’s Bay. Amongst all sites, Kalk Bay had the highest annual maximum temperature of 19 ℃. Annual minimum coastal water temperatures for the study sites on the western side of the peninsula ranged from 12.9 ℃ at St. Helena Bay to 14.1 ℃ at Olifantsbos, and on the western side of the peninsula ranged from 15.0 ℃ at Buffelsbaai to 13.6 ℃ at Betty’s Bay.  The annual range in temperature on the western side on the peninsula was 3.4 at St. Helena Bay to 0.9 at Olifantsbos, and on the eastern side on of the peninsula was 1.0 at Buffelsbaai to 2.9 at Betty’s Bay.

@@

temp_boxplot <- ggplot(data = temp_ann_meta, aes(x = region, y = mean_temp, fill  = region)) + 
  geom_boxplot(show.legend = FALSE, width = 0.5) + 
  stat_compare_means(label.x = 2.5, label.y = 12.5, label = "p.signif") +
  xlab("") + 
  ylab("") +
  coord_flip() + 
  theme_bw() +
  scale_fill_grey(start = 0.4, end = 0.8) + 
  theme(axis.text.y=element_text(vjust=0.3))

@@

In general, the mean temperature increases for sites located within False Bay with less variation around the mean temperature compared to sites on the western side of the peninsula(\ref{fig:Temperature variables plot}A). This is verified by figure\ref{fig:Temperature variables plot}B, which shows a significantly lower (*p* < 0.05, figure \ref{fig:Temperature variables plot}B) mean temperature and less variation in temperature for sites located on the western side of the peninsula compared to sites on the eastern side.

@@

Our data show the western side of the peninsula experiences higher significant wave heights and variation in wave heights compared to the eastern side of the peninsula on both an annual scale (see figure \ref{fig:wave variable plot}). Mean significant wave height ranged from 0.5 ± 0.2 m (mean ± SD) at St. Helena Bay to 2.3  ± 0.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side of the peninsula it ranged from 0.9 ± 0.4 m at Buffelsbaai to 1.9 ± 0.6 m at Betty’s Bay (see figure \ref{fig:wave variable plot}A). The maximum significant wave height ranged from 1.4 m at St. Helena Bay to 4.7 m at Olifantsbos on the western side of the peninsula, and on the eastern side ranged from 2.6 m at Buffelsbaai to 4.2 m at Betty’s Bay. Mean peak period for sites on the western side of the peninsula ranged from 10.3 ± 3.5 s at St. Helena Bay to 11.0 ± 2.0 s at Olifantsbos, and ranged from 10.6 ± 3.0 s at Buffelsbaai to 10.8 ± 2.2 s at Betty’s Bay (see figure \ref{fig:wave variable plot}B). The data show no trend in mean peak period for the coastline. These data show that the western side of the peninsula has a lower variation (SD) compared to the eastern side, with Miller’s Point showing the highest variation across timescales. Maximum peak period ranged from 18.9 s at St. Helena Bay to 18.4 s at Olifantsbos for the western side of the peninsula, and ranged from 18.0 s at Buffelsbaai to 18.0 s at Betty’s Bay on the eastern side of the peninsula. When wave variables were grouped by region significant differences were found for both mean significant swell height and mean peak period. The western region experiences a significantly higher mean significant swell height (*p* < 0.05; see figure \ref{fig:wave variable plot}C) and mean peak period (*p* < 0.05; see figure \ref{fig:wave variable plot}D).

@@

wave_boxplot_1 <- ggplot(data = wave_ann_meta, aes(x = region, y = hs_mean, fill  = region)) + 
  geom_boxplot(show.legend = FALSE) + 
  stat_compare_means(label.y =0.6 ,label.x = 2.5, label = "p.signif") + 
  xlab("") + 
  ylab("") + 
  coord_flip() + 
  theme_bw() +
  scale_fill_grey(start = 0.4, end = 0.8) + 
  theme(axis.text.y=element_text(vjust=0.3))

wave_boxplot_2 <- ggplot(data = wave_ann_meta, aes(x = region, y = tp_mean, fill  = region)) + 
  geom_boxplot(show.legend = FALSE) + 
  stat_compare_means(label.y = 8.2, label.x = 2.5, label = "p.signif") + 
  xlab("") + 
  ylab("") +
  ylim(8,12) + 
  coord_flip() + 
  theme_bw() +
  scale_fill_grey(start = 0.4, end = 0.8) + 
  theme(axis.text.y=element_text(vjust=0.3))

@@

In figure \ref{fig:wave power map} the total coastal wave exposure of the Cape Peninsula is given in terms of wave energy (kW per meter wave crest length). The classification from fully sheltered to extremely exposed is based on the total wave energy upper and lower limits. The western periphery of the Cape Peninsula is almost continuously exposed to high wave exposures while the eastern periphery of the peninsula (western coastline of False Bay) revealed sheltered wave exposures (see figure \ref{fig:wave power map}). There the marked seasonality, with higher energy waves during winter, may be clearly observed once more.

@@

#temperature and lamanaria  
LT1 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaWeight, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.62, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina weight") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT2 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina length") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT3 <- ggplot(data = lam_data, aes(x = temp_regime, y = laminaThickness, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Lamina thickness") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT4 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT5 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe length (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT6 <- ggplot(data = lam_data, aes(x = temp_regime, y = stipeDiameter, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Stipe diameter (cm)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT7 <- ggplot(data = lam_data, aes(x = temp_regime, y = digits, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("No. of digits") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT8 <- ggplot(data = lam_data, aes(x = temp_regime, y = thallusMass, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.6, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Thallus mass (g)") +
  scale_fill_discrete(name="Temperature regime") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

LT9 <- ggplot(data = lam_data, aes(x = temp_regime, y = totalLength, fill = temp_regime)) +
  scale_x_discrete(limits = pos_temp) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = FALSE, show.legend = FALSE) + 
  stat_compare_means(method = "wilcox.test", label = "p.signif", label.x  = 0.65, label.y.npc = c('top')) + 
  xlab("") + 
  ylab("Total length (cm)") +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("") +
  scale_fill_grey(start = 0.8, end = 0.4)

@@

LW1 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaWeight, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.85, label.y.npc = c('top')) + 
  ylab("Lamina weight (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW2 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("Lamina length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW3 <- ggplot(data = lam_data, aes(x = wave_exposure, y = laminaThickness, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam) + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("Lamina thickness") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW4 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeMass, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe mass (g)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW5 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeLength, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe length (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW6 <- ggplot(data = lam_data, aes(x = wave_exposure, y = stipeDiameter, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.9, label.y.npc = c('top')) + 
  ylab("Stipe diameter (cm)") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")

LW7 <- ggplot(data = lam_data, aes(x = wave_exposure, y = digits, fill = wave_exposure)) +
  scale_x_discrete(limits = pos_exposure_lam)  + 
  geom_boxplot(outlier.shape = NA, varwidth = TRUE, notch = TRUE) +
  stat_compare_means(method = "kruskal.test", label = "p.signif", label.x  = 0.82, label.y.npc = c('top')) + 
  ylab("No. of digits") + 
  scale_fill_brewer(palette="OrRd") +
  guides(fill=guide_legend(title="Wave exposure level")) +
  theme_bw() +
  theme(axis.text.x = element_blank()) + 
  xlab("")
